---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
devtools::load_all(".")
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# ucdplink

The goal of **ucdplink** is to link data on violent events from the Uppsala Conflict Data Program (UCDP) Georeferenced Event Dataset (GED) to armed conflicts listed in the UCDP Armed Conflict Dataset (ACD).

The rationale for this procedure is described in the article: Bara, Corinne & Maurice P. Schumann, "Who, what, and where? Linking violence to civil wars", *Research & Politics* (2025).

Please cite the article when using this data or package.

**The logic in brief:** Civil wars are more than battles between governments and rebels; they involve a multitude of actors who perpetrate different forms of violence linked to the war in some way. In practice, however, linking various forms of violence to one another and attributing them to a specific civil war remains a challenge, conceptually and practically. The ucdplink package solves the practical aspects of the violence attribution problem and links casualty counts from different forms of war-related violence to specific civil wars using information on the actors involved in the event, and on the event location.

The main function **ucdplink** produces a conflict-month dataset with information on monthly casualties from various forms of violence. For more details on the variables included in the final dataset please consult the [Codebook](https://github.com/corinnebara/ucdplink/releases/download/v1.0.0/Codebook.pdf). The procedure has three steps:

1.  It downloads the UCDP/PRIO Armed Conflict Dataset (ACD) and turns it from annual data into a conflict-month dataset. This is done in the function **acdtomonthly**, which may be useful for users even if they are not interested in violence linking.

2.  It then downloads the UCDP Georeferenced Event Dataset (GED) and performs some common data wrangling steps, namely splitting multi-month events into separate monthly events with casualty counts evenly divided between them, and splitting events perpetrated by coalition actors among the coalition-constituting actors. This is done in the function **gedprep,** which may also be useful even for users not interested in violence linking.

3.  Finally, the function links casualties from different forms of violence (state-based, one-sided, non-state) to the conflict episodes in which this violence (most likely) takes place. The linking is done via the actors involved in the violence, and - if that is not possible - via the location of the event, using information on the location of conflicts from the WZONE Conflict Zones Data.

The package currently processes UCDP data versions 24.1. When using this package, make sure you also correctly cite the source data.

For the function **acdtomonthly**: Cite UCDP/PRIO Armed Conflict Dataset v.24.1. For citation details, see [UCDP Dataset Download Center](https://ucdp.uu.se/downloads/).

For the function **gedprep**: Cite UCDP Georeferenced Event Dataset (GED) Global version 24.1. For citation details, see [UCDP Dataset Download Center](https://ucdp.uu.se/downloads/).

For the function **ucdplink**: Cite both the above, and in addition the WZONE conflict zone data: [WZONE: Zones of Armed Conflicts - Harvard Dataverse](https://dataverse.harvard.edu/dataset.xhtml?persistentId=doi:10.7910/DVN/PUWJEU).

## Installation

You can install the package from github as follows:

``` r
library(devtools)

devtools::install_github("corinnebara/ucdplink")
library(ucdplink)
```

You also need the `dplyr`, `tidyr`, `stringr`, `readr`, `lubridate`, `sf`, `cshapes`, `future`, and `furrr` packages. R will try to install these dependencies if not already installed.

## The Final Data

**Note:** Running the `ucdplink` function may take a considerable amount of time due to computationally intensive spatial operations. Users who are happy with our default choices (see below) can also simply download the final dataset. It includes conflicts from 1989 to 2023 while they are active, and includes all inactive months until the end of 2023. Users can simply filter the amount of postwar months they want included for each conflict episode via the *postwarmonth* variable.

-   [Final Dataset Excel](https://github.com/corinnebara/ucdplink/releases/download/v1.0.0/Final_dataset.xlsx)

-   [Final Dataset R Data](https://github.com/corinnebara/ucdplink/releases/download/v1.0.0/Final_dataset.rds)

## ucdplink

The function takes the following arguments:

-   postwar_months: How many inactive months to add to each episode. Default is 0 (only active conflict). Add 5000 if you want all possible postwar months until the end of the dataset.

-   include_alqaida: Whether to include conflict 418 (US vs. al-Qaida), default is to exclude, as this conflict did not take place on US territory.

-   start_year: Specify the start year. Default: 1989.

-   end_year: Specify the end year. Default: longest possible (currently 2023).

-   buffer_percent: How much buffer to add around each conflict zone (in percent, numeric from 0 to 100). Default is no buffer.

-   clipcountry: Whether to clip conflict zones at the borders of the country/countries primarily affected by the conflict. Default is to clip.

-   divide_deaths: For events that fall into multiple conflict zones: use full casualties in each or divide deaths between the conflicts? Default is to use full.

-   include_gedtrack: Whether to also return a version of GED in which event has information on how it was linked to the conflict-month dataset. Default is to not include.

Please refer to Bara & Schumann (2025) for a discussion on the implications of different user choices.

```{r example1, eval = FALSE}
linkeddata <- ucdplink(postwar_months = 0, include_alqaida = FALSE, start_year = 1989, end_year = 2023, buffer_percent = 0, clipcountry = TRUE, divide_deaths = FALSE, include_gedtrack = FALSE)
```

## acdtomonthly

The function takes the following arguments:

-   postwar_months: How many inactive months to add to each episode. Default is 0 (only active conflict). Add 5000 if you want all possible postwar months until the end of the dataset.

-   include_alqaida: Whether to include conflict 418 (US vs. al-Qaida), default is to exclude, as this conflict did not take place on US territory.

-   start_year: Specify the start year. Default: 1989.

-   end_year: Specify the end year. Default: longest possible (currently 2023).

```{r example2, eval = FALSE}
monthlydata <- acdtomonthly(postwar_months = 0, include_alqaida = FALSE, start_year = 1989, end_year = 2023)
```

## gedprep

The function takes no arguments:

```{r example3, eval = FALSE}
preppedged <- gedprep()
```

## Work in Progress

-   In the future, we plan to allow users to choose the UCDP version they want to work with. As of now, the package only works with the (currently) newest version of UCDP, i.e., v24.

-   We also plan to implement an additional user choice to only link those forms of violence that can be linked via actor identifiers (battle, rebel one-sided violence, and inter-rebel fighting), and forego the more time-consuming spatial linking.
